---
title: ""
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(dplyr)
library(DT)
library(ggplot2)
library(googlesheets4)
library(here)
library(htmltools)
library(knitr)
library(leaflet)
library(lubridate)
library(plotly)
library(purrr)
library(stringr)
library(summaryplots)
library(tidyr)
library(waves)

source(here("functions/helpers.R"))
source(here("functions/subchunkify.R"))

theme_set(theme_light())

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  scrollX = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

zero_crossing <- c(
  "average_height_largest_33_percent_m",
  "average_height_largest_10_percent_m",
  "period_largest_33_percent_s",
  "period_largest_10_percent_s",
  "maximum_height_m",
  "period_maximum_s"
)

# read in all wave data
dat <- readRDS(here("data/2024-08-28_wave_data_prelim_qc.rds")) %>% 
  filter((
    !variable %in% zero_crossing),
    deployment_id != "PC001"
  )

dat_wide <- dat %>% 
  pivot_wider(values_from = "value", names_from = "variable")


gs4_deauth()
st_locations <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1DVfJbraoWL-BW8-Aiypz8GZh1sDS6-HtYCSrnOSW07U/edit?gid=1923344665#gid=1923344665",
  sheet = "Area Info",
  col_types = "cccnnc", na = c("", "NA")
) %>% 
  inner_join(distinct(dat, county, station), by = join_by(station, county))


county_pal <- get_county_colour_palette(length(unique(dat$county)))

theme_facet_plotly <- theme(
  panel.spacing.y = unit(15, "lines"),
  panel.spacing.x = unit(15, "lines")
)

vars <- unique(dat$variable)

```

# Wave Parameters: Explore Peak Period

`r Sys.Date()`

- Review observations where peak_period_s = 25.6 seconds
- Review observations where peak_period_s > 18 seconds
- Preliminary QC applied (0 values and obvious outliers removed through grossrange test).
- Additional QC applied after additional review of the data
  - Wave heights > 20 m
  - Periods > 100 s

# Station Locations

Approximate location of stations with wave data.

```{r}
#| fig-height: 6

county_map_pal <- colorFactor(
  palette = county_pal,
  domain = unique(st_locations$county),
  na.color = "transparent"
)

# interactive map
leaflet(st_locations) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircleMarkers(
    data = st_locations,
    lng = ~longitude, lat = ~latitude, label = ~station,
    weight = 1,
    color = "black",
    fillOpacity = 0.75,
    fillColor = ~county_map_pal(county),
    radius = 5
  ) %>% 
  addScaleBar(
    position = "bottomleft", 
    options = scaleBarOptions(imperial = FALSE)
  )

```

# Period = 25.6 seconds

### Table 1
Deployments with peak_period_s = 25.6
```{r}
#| message: false

period_25 <- dat_wide %>% 
  filter(peak_period_s == 25.6)

# deployments with period = 25.6
period_25 %>% 
  group_by(county, waterbody, station, deployment_id) %>% 
  arrange(county) %>% 
  summarise(n = n()) %>% 
  datatable(options = dt_options, rownames = FALSE)
```

### Figure 1
Approximate date of when peak_period = 25.6 
```{r}
#| fig-height: 10

t <- period_25 %>% 
  mutate(date_utc = as_date(timestamp_utc)) %>% 
  distinct(county, waterbody, station, deployment_id, date_utc)

t2 <- period_25 %>% 
  mutate(
    year_utc = year(timestamp_utc),
    month_utc = month(timestamp_utc),
  ) %>% 
  distinct(county, waterbody, station, deployment_id, year_utc, month_utc)

p <- ggplot(
  t, 
  aes(date_utc, fill = deployment_id,
      text = paste0("depl_id: ", deployment_id, "\n",
                    "date: ", as.character(date_utc))
  )
) +
  geom_histogram(bins = 30) +
  facet_wrap(~county, ncol = 3, scales = "free_y") +
  theme_facet_plotly

ggplotly(p, tooltip = "text")
```

## Figure 2

Significant wave height recorded when peak_period_s = 25.6
```{r}
#| fig-height: 10

x <- period_25 %>% 
  filter(significant_height_m > 0.25)


p1 <- ggplot(period_25, aes(significant_height_m)) +
  geom_histogram(binwidth = 0.1)
  
p <- ggplot(
  period_25, 
  aes(significant_height_m, fill = deployment_id,
      text = paste0("depl_id: ", deployment_id, "\n",
                    "height: ", significant_height_m))
) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~county, ncol = 3, scales = "free_y") +
  theme_facet_plotly

ggplotly(p, tooltip = "text")

```

# Significant wave heights > 0.25 m 

### Table 2
Significant wave heights > 0.25 m recorded when peak_period_s = 25.6
```{r}
#| message: false
period_25_large <- period_25 %>% 
  filter(significant_height_m > 0.25) 

period_25_large %>%  
  group_by(county, waterbody, station, deployment_id) %>% 
  arrange(county) %>% 
  summarise(n = n()) %>% 
  datatable(options = dt_options, rownames = FALSE)
```

### Figure 3
```{r}
#| fig-height: 10

p0 <- ggplot(period_25_large, aes(significant_height_m)) +
  geom_histogram(binwidth = 0.05)


p <- ggplot(
  period_25_large, 
  aes(significant_height_m, fill = deployment_id,
      text = paste0("depl_id: ", deployment_id, "\n",
                    "height: ", significant_height_m))
) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~county, ncol = 3, scales = "free_y") +
  theme_facet_plotly

ggplotly(p, tooltip = "text")
```


# Period > 25.6 seconds

Deployments with peak_period_s > 25.6

### Table 3

```{r}

period_26 <- dat_wide %>% 
  filter(peak_period_s > 25.6)

# deployments with period = 25.6
period_26 %>% 
  distinct(county, waterbody, station, deployment_id) %>% 
  arrange(county) %>% 
  datatable(options = dt_options, rownames = FALSE)
```

### Figure 4
Approximate date when peak_period > 25.6 
```{r}
#| fig-height: 10

t <- period_26 %>% 
  mutate(date_utc = as_date(timestamp_utc)) %>% 
  distinct(county, waterbody, station, deployment_id, date_utc)

p <- ggplot(
  t, 
  aes(date_utc, fill = deployment_id,
      text = paste0("depl_id: ", deployment_id, "\n",
                    "date: ", as.character(date_utc))
  )
) +
  geom_histogram(bins = 30) +
  facet_wrap(~county, ncol = 3, scales = "free_y") +
  theme_facet_plotly

ggplotly(p, tooltip = "text")
```

## Figure 5

Significant wave height recorded when peak_period_s > 25.6
```{r}
#| fig-height: 10

p <- ggplot(
  period_26, 
  aes(significant_height_m, fill = deployment_id,
      text = paste0("depl_id: ", deployment_id, "\n",
                    "height: ", significant_height_m))
) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~county, ncol = 3, scales = "free_y") +
  theme_facet_plotly

ggplotly(p, tooltip = "text")

```


# Period > 18 seconds (not including period = 25.6 s)

### Table 4
Deployments with peak_period_s > 18 & peak_period_s != 25.6
```{r}

period_18 <- dat_wide %>% 
  filter(peak_period_s > 18 & peak_period_s != 25.6)

# deployments with period = 25.6
period_18 %>% 
  distinct(county, waterbody, station, deployment_id) %>% 
  arrange(county) %>% 
  datatable(options = dt_options, rownames = FALSE)
```

### Figure 6
Approximate date of when peak_period > 18 
```{r}
#| fig-height: 10

t <- period_18 %>% 
  mutate(date_utc = as_date(timestamp_utc)) %>% 
  distinct(county, waterbody, station, deployment_id, date_utc)

p <- ggplot(
  t, 
  aes(date_utc, fill = deployment_id,
      text = paste0("depl_id: ", deployment_id, "\n",
                    "date: ", as.character(date_utc))
  )
) +
  geom_histogram(bins = 30) +
  facet_wrap(~county, ncol = 3, scales = "free_y") +
  theme_facet_plotly

ggplotly(p, tooltip = "text")
```

## Figure 7

Significant wave height recorded when peak_period_s > 25.6
```{r}
#| fig-height: 10

p <- ggplot(
  period_18, 
  aes(significant_height_m, fill = deployment_id,
      text = paste0("depl_id: ", deployment_id, "\n",
                    "height: ", significant_height_m))
) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~county, ncol = 3, scales = "free_y") +
  theme_facet_plotly

ggplotly(p, tooltip = "text")

```
