---
title: ""
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(dplyr)
library(DT)
library(ggplot2)
library(googlesheets4)
library(here)
library(htmltools)
library(knitr)
library(leaflet)
library(lubridate)
library(plotly)
library(purrr)
library(stringr)
library(summaryplots)
library(tidyr)
library(waves)

source(here("functions/helpers.R"))
source(here("functions/subchunkify.R"))

theme_set(theme_light())

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  scrollX = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

zero_crossing <- c(
  "average_height_largest_33_percent_m",
  "average_height_largest_10_percent_m",
  "period_largest_33_percent_s",
  "period_largest_10_percent_s",
  "maximum_height_m",
  "period_maximum_s"
)

# read in all wave data
dat <- readRDS(here("data/2024-08-28_wave_data_no_qc.rds")) %>% 
  filter((!variable %in% zero_crossing))

dat_wide <- dat %>% 
  pivot_wider(values_from = "value", names_from = "variable")

link <- "https://docs.google.com/spreadsheets/d/1DVfJbraoWL-BW8-Aiypz8GZh1sDS6-HtYCSrnOSW07U/edit?gid=1923344665#gid=1923344665"
gs4_deauth()
st_locations <- read_sheet(
  link,
  sheet = "Area Info",
  col_types = "cccnnc", na = c("", "NA")
) %>% 
  inner_join(
    distinct(dat, county, station, deployment_id),
    by = join_by(station, county), 
    relationship = "many-to-many") %>% 
  mutate(label = paste(station, deployment_id, sep = "</br>"))

county_pal <- get_county_colour_palette(length(unique(dat$county)))

theme_facet_plotly <- theme(
  panel.spacing.y = unit(15, "lines"),
  panel.spacing.x = unit(15, "lines")
)

```

# Wave Parameters: Explore Peak Period

`r Sys.Date()`

- No preliminary QC applied

# Station Locations

Approximate location of stations with wave data.

```{r}
#| fig-height: 6

county_map_pal <- colorFactor(
  palette = county_pal,
  domain = unique(st_locations$county),
  na.color = "transparent"
)

# interactive map
leaflet(st_locations) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addCircleMarkers(
    data = st_locations,
    lng = ~longitude, lat = ~latitude, popup = ~label,
    weight = 1,
    color = "black",
    fillOpacity = 0.75,
    fillColor = ~county_map_pal(county),
    radius = 5
  ) %>% 
  addScaleBar(
    position = "bottomleft", 
    options = scaleBarOptions(imperial = FALSE)
  )

```

# Period = 25.6 seconds

:::panel-tabset
### Figure 1

Significant wave height recorded when peak_period_s = 25.6
```{r}
#| fig-height: 10

period_25 <- dat_wide %>% 
  filter(peak_period_s == 25.6)
  
p <- ggplot(
  period_25, 
  aes(significant_height_m, fill = deployment_id,
      text = paste0("depl_id: ", deployment_id, "\n",
                    "height: ", significant_height_m))
) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~county, ncol = 3, scales = "free_y") +
  theme_facet_plotly

ggplotly(p, tooltip = "text")

```

### Table 1
Deployments with peak_period_s = 25.6
```{r}
#| message: false

# deployments with period = 25.6
period_25_depls <- period_25 %>% 
  group_by(county, waterbody, station, deployment_id) %>% 
  arrange(county) %>% 
  summarise(n = n()) 

datatable(period_25_depls, options = dt_options, rownames = TRUE)
```

:::


# Significant wave heights > 0.25 m 

::: panel-tabset

### Figure 2
```{r}
#| fig-height: 10
#| warning: false

period_25_large <- period_25 %>% 
  filter(significant_height_m > 0.25) 

p <- ggplot(
  period_25_large, 
  aes(significant_height_m, fill = deployment_id,
      text = paste0("depl_id: ", deployment_id, "\n",
                    "height: ", significant_height_m))
) +
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(limits = c(0, 2)) +
  facet_wrap(~county, ncol = 3, scales = "free_y") +
  theme_facet_plotly

ggplotly(p, tooltip = "text")
```

### Table 2
Significant wave heights > 0.25 m recorded when peak_period_s = 25.6
```{r}
#| message: false

period_25_large_depls <- period_25_large %>%  
  group_by(county, waterbody, station, deployment_id) %>% 
  arrange(county) %>% 
  summarise(n = n()) 

datatable(period_25_large_depls, options = dt_options, rownames = TRUE)
```
:::

# Deployment Figures

## Deployments with > 100 observations of Tp = 25.6 s

::: panel-tabset
```{r}
#| warning: false
#| message: false
#| results: asis

depls <- period_25_depls %>% 
  arrange(n) %>% 
  filter(n > 100)
depls <- depls$deployment_id

vars <- c("significant_height_m", "peak_period_s", "to_direction_degree")   

for(i in seq_along(depls)) {

  depl_i <- depls[i]

  dat_i <- dat %>%
    filter(deployment_id == depl_i, variable %in% vars)

  cat(paste("### ", depl_i, "\n \n"))

  p_i <- dat_i %>%
    wv_plot_ts(scales = "free")

  subchunkify(p_i, fig_height = 8, fig_width = 8.5)

  cat("\n\n")
}

```
:::

## Deployments with > 5 observations of Hs > 0.25 m

::: panel-tabset
```{r}
#| warning: false
#| message: false
#| results: asis

depls <- period_25_large_depls %>% 
  arrange(n) %>% 
  filter(n > 5)
depls <- depls$deployment_id

for(i in seq_along(depls)) {

  depl_i <- depls[i]

  dat_i <- dat %>%
    filter(deployment_id == depl_i, variable %in% vars)

  cat(paste("### ", depl_i, "\n \n"))

  p_i <- dat_i %>%
    wv_plot_ts(scales = "free")

  subchunkify(p_i, fig_height = 8, fig_width = 8.5)

  cat("\n\n")
}

```
:::

## GY006

::: panel-tabset
```{r}
#| warning: false
#| message: false
#| results: asis

dat_i <- dat %>%
  filter(deployment_id == "GY006", variable %in% vars)

cat(paste("### ", depl_i, "\n \n"))

p_i <- dat_i %>%
  wv_plot_ts(scales = "free")

subchunkify(p_i, fig_height = 8, fig_width = 8.5)

```
:::

