---
title: ""
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(dplyr)
library(DT)
library(ggplot2)
library(googlesheets4)
library(here)
library(htmltools)
library(knitr)
library(leaflet)
library(plotly)
library(purrr)
library(stringr)
library(summaryplots)
library(tidyr)
library(waves)
library(zoo)

source(here("functions/helpers.R"))
source(here("functions/subchunkify.R"))

theme_set(theme_light())

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  scrollX = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

zero_crossing <- c(
  "average_height_largest_33_percent_m",
  "average_height_largest_10_percent_m",
  "period_largest_33_percent_s",
  "period_largest_10_percent_s",
  "maximum_height_m",
  "period_maximum_s"
)

# read in all wave data
dat <- readRDS(here("data/2024-08-28_wave_data_prelim_qc.rds")) %>% 
  filter(
    !(variable == "average_height_largest_33_percent_m" & value > 15),
    !(variable == "period_largest_10_percent_s" & value > 60),
    !(variable == "period_largest_33_percent_s" & value > 60),
    !(variable == "period_maximum_s" & value > 60),
    !(variable %in% zero_crossing & deployment_id == "IV001"),
    !(variable %in% zero_crossing & deployment_id == "QN010"),
    deployment_id != "PC001"
  )

gs4_deauth()
st_locations <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1DVfJbraoWL-BW8-Aiypz8GZh1sDS6-HtYCSrnOSW07U/edit?gid=1923344665#gid=1923344665",
  sheet = "Area Info",
  col_types = "cccnnc", na = c("", "NA")
) %>%
  inner_join(distinct(dat, county, station), by = join_by(station, county))


county_pal <- get_county_colour_palette(length(unique(dat$county)))

theme_facet_plotly <- theme(panel.spacing.y = unit(1, "lines"))

vars <- unique(dat$variable)

period_hours <- 12
max_interval_hours <- 4

```

# Wave Parameters: Explore rolling standard deviation

`r Sys.Date()`

- Preliminary QC applied (0 values and obvious outliers removed through grossrange test).
- Additional QC applied after additional review of the data
  - Wave heights > 20 m
  - Periods > 100 s
- Removed zero-crossing variables from IV001.
- Removed all of PC001.

# Station Locations

Approximate location of stations with wave data.

```{r}
#| fig-height: 6

county_map_pal <- colorFactor(
  palette = county_pal,
  domain = unique(st_locations$county),
  na.color = "transparent"
)

# interactive map
leaflet(st_locations) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    data = st_locations,
    lng = ~longitude, lat = ~latitude, label = ~station,
    weight = 1,
    color = "black",
    fillOpacity = 0.75,
    fillColor = ~county_map_pal(county),
    radius = 5
  ) %>%
  addScaleBar(
    position = "bottomleft",
    options = scaleBarOptions(imperial = FALSE)
  )

```

# Sample Interval

- Interval (minutes) between consecutive observations.
- Expect 60 minutes between most samples.
- Some QC has been applied to the observations, resulting in large gaps between some observations.

```{r}
dat_int <- dat %>%
  group_by(county, station, deployment_id, variable) %>%
  dplyr::arrange(timestamp_utc, .by_group = TRUE) %>%
  mutate(
    # sample interval
    int_sample = difftime(timestamp_utc, lag(timestamp_utc), units = "mins"),
    int_sample = round(as.numeric(int_sample))
  ) 
```

::: panel-tabset

## Figure 1
```{r}
#| message: false
#| warning: false
#| fig-height: 10

dat_int %>% 
  na.omit() %>% 
  wv_convert_vars_to_ordered_factor() %>% 
  plot_histogram(hist_col = int_sample, binwidth = 500) +
  facet_wrap(~variable, scales = "free", ncol = 2)

```

## Figure 2

- Filter so `int_sample` < 240 mins
- Most of the remaining observations have an interval of 60 mins
```{r}
#| message: false
#| warning: false
#| fig-height: 10

dat_int %>% 
  filter(int_sample < 240) %>% 
  na.omit() %>% 
  wv_convert_vars_to_ordered_factor() %>% 
  plot_histogram(hist_col = int_sample, binwidth = 10) +
  facet_wrap(~variable, scales = "free", ncol = 2)
```

:::

# Number of Samples Per Period

- Filtered to only include samples with time interval < 240 min (4 hours)
- Expect 12 samples in 12 hours
- `period_hours` = `r period_hours`
- `max_interval_hours` = `r max_interval_hours`

```{r}

dat_sample <- dat_int %>%
  filter(int_sample < 240) %>% # not sure if this matters
  mutate(
    # number of samples in  period_hours
    # 60 mins / hour * 1 sample / int_sample mins * 24 hours / period
    n_sample = round((60 / int_sample) * period_hours),
    # n_sample = if_else(is.na(n_sample), 1, n_sample), # first obs
    n_sample_effective = case_when(
      # first observation of each group is NA, which will give error in rollapply
      is.na(n_sample) ~ 0,
      # if the sample interval is greater than acceptable limit, set n_sample to 0
      # so that roll_sd will be NA
      int_sample > max_interval_hours * 60 ~ 0,
      TRUE ~ n_sample
    ))
```

::: panel-tabset
## Figure 3

- n_sample

```{r}
#| message: false
#| warning: false
#| fig-height: 10

plot_histogram(dat_sample, hist_col = n_sample) +
  facet_wrap(~variable, scales = "free", ncol = 2)
```

## Figure 4

- n_sample_effective

```{r}
#| message: false
#| warning: false
#| fig-height: 10

plot_histogram(dat_sample, hist_col = n_sample_effective) +
  facet_wrap(~variable, scales = "free", ncol = 2)
```
:::

## Standard Deviation

```{r}

dat_sd <- dat_sample %>%
  mutate(
    # rolling sd
    sd_roll = rollapply(
      value,
      width = n_sample_effective,
      align = "center", FUN = sd, fill = NA
    ),
    sd_roll = round(sd_roll, digits = 2)
  )
```

::: panel-tabset
## Figure 5
```{r}
#| message: false
#| warning: false
#| fig-height: 10

plot_histogram(dat_sd, hist_col = sd_roll) +
  facet_wrap(~variable, ncol = 2, scales = "free")
```

## Figure 6

- Only use observations with n_sample_effective = 12

```{r}
#| message: false
#| warning: false
#| fig-height: 10

dat_sd %>% 
  filter(n_sample_effective == 12) %>% 
  plot_histogram(hist_col = sd_roll) +
  facet_wrap(~variable, ncol = 2, scales = "free")
```
:::
